---
title: "Profolio2"
author: "Wei-Tang, Chen"
date: "21/05/2018"
output: html_document
---

```{r setup, include=FALSE}
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)
```

Task1.1 
In what day of the week are the largest number of Uber pick-ups? and the minimum? Perform this analysis both calculating statistics from the dataset and visualizing the data to show daily trends in pick-ups. 
```{r}
#install.packages("ggplot2")
#install.packages("reshape2")
#install.packages("ggmap")
#install.packages("plyr")
#install.packages("grid")
#Load the data
Data1 <-read.csv("uber-raw-data-apr14.csv", sep=",", header = T, na.strings = NA)
Data2 <-read.csv("uber-raw-data-may14.csv", sep=",", header = T, na.strings = NA)
Data3 <-read.csv("uber-raw-data-jun14.csv", sep=",", header = T, na.strings = NA)
Data4 <-read.csv("uber-raw-data-jul14.csv", sep=",", header = T, na.strings = NA)
Data5 <-read.csv("uber-raw-data-aug14.csv", sep=",", header = T, na.strings = NA)
Data6 <-read.csv("uber-raw-data-sep14.csv", sep=",", header = T, na.strings = NA)
#Combine the datainto one
uberRawData <- rbind(Data1,Data2,Data3,Data4,Data5,Data6)
uberRawData$Date.Time <- as.character(uberRawData$Date.Time)
uberRawData$Base <- as.character(uberRawData$Base)
#split the column by / to extract the month, day and time
monthAndDay <- strsplit(uberRawData$Date.Time, "/")
#Use do.call to split the one column of the list
monthAndDay <- do.call(rbind, monthAndDay)
colnames(monthAndDay) <- c("month", "day", "Time")
monthAndDay <- data.frame(monthAndDay)
#You need to use as.character before as.numeric. This is because factors are stored internally as integers with a table to give the factor level labels. Just using as.numeric will only give the internal integer codes. 
monthAndDay$month <- as.numeric(as.character(monthAndDay$month))
monthAndDay$day <- as.numeric(as.character(monthAndDay$day))
#use order to arrange the element by specific column
#Note the trailing comma; what this means is to take all the columns.  If we only wanted certain columns, we could specify after this comma.
monthAndDay <- monthAndDay[order(monthAndDay$month),]
#use table and it will calculate the frequency automatically
date <- table(monthAndDay$month, monthAndDay$day)
#create a table like "Calender""
vec_mix <- c(date[1,-c(ncol(date))], date[2,], date[3,-c(ncol(date))], date[4,], date[5,], date[6,-c(ncol(date))])
#To fill the rest empty element in order to create a matrix below
vec_mix[184:196] <- 0
#Create a calender with the name of weekday
mat_all1 <- matrix(vec_mix, ncol=7, byrow = T)
mat_all2 <- mat_all1
colnames(mat_all2) <- c("Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday", "Monday")
a=1
#extract column name
col_mat <- colnames(mat_all2)
#initialize the vector "total""
total<-c()
#Calculate the sum of each day
for(i in 1:7){
  total[i] <- sum(mat_all2[,i])
}
daily_list <- cbind(col_mat,total)
daily_list <- data.frame(daily_list)
#(use as.data.frame to calculate the frequency as well)
library(ggplot2)
library(reshape2)
#Reorder the levels of factor
daily_list$col_mat <-factor(daily_list$col_mat, levels =  c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))
#Plot the bar plot
ggplot(daily_list, aes(x=col_mat, y=total, fill="red")) + geom_bar(stat = "identity", colour="black")+labs(title="Daily Trend", subtitle="From Monday to Sunday", y="Pick-up times", x="Weekday")+theme(legend.position = "None", plot.title = element_text(size=20, face="bold", color="tomato"))
#Summary: Thursday has the largest number of Uber pick-ups, while Sunday has least.
#-----------------------------------------------------------------------------------
```
Task 1.2  How did the use of Uber in NYC trend over the months considered in the dataset? Does it increase? Perform this analysis through visualization.
```{r}
library(grid)
#change the type into character in order to use the labels
monthAndDay$month <- as.character(monthAndDay$month)
#use ggplot to count(stat->count) the frequency of the pick-ups through the month, use the breaks to customize the ticks of x axis
gg<-ggplot(monthAndDay, aes(x=monthAndDay$month)) + geom_bar(stat="count") + scale_x_discrete(breaks=seq(4,9,1),labels=c("April","May","June","July","August","September"))
#print out
my_text <- "There is a raising trend from April to September!"
my_grob = grid.text(my_text, x=0.3,  y=0.8, gp=gpar(col="firebrick", fontsize=10, fontface="bold"))

gg+labs(title="Monthly Trend", subtitle="From April to September", y="Pick-up times", x="Month")+theme(panel.background = element_rect((fill="khaki")),plot.title = element_text(size=20, face="bold", color="black"),plot.background=element_rect(fill="salmon"))+annotation_custom(my_grob)

#Summary THe Uber pick-ups increased by the months according to the statistical graph.
```
Task 1.3 How do pick-up patterns change over time of day? What times of the day are the most popular in terms of pick-ups? Perform this analysis through visualization.
```{r}
#extract tje data and make it a formal time format in order to cut the specific column
time_day <- data.frame((monthAndDay$month),substring(as.POSIXct(monthAndDay$Time, format="%Y %H:%M:%S"), 12,13 ))
#create table tp get the frequency of pick-ups
table1 <- table(time_day)
#make it a data frame for ggplot
table1 <- data.frame(table1)
colnames(table1) <- c("Month","Time","Freq")
#cancel the scientific notation
ggplot(table1, aes(x=Time, y=Freq, fill=Month))+geom_bar(stat = "identity")+scale_y_continuous(labels = function(x) format(x, scientific = F))+labs(title="Hourly Trend", subtitle="From 00:00AM to 23:59PM", y="Pick-up times", x="Hour")+theme(,plot.title = element_text(size=20, face="bold", color="black"))
#Summary : In the morning the pick-ups are apprently much lower than the afternoon(15:00) to evening(22:00) .
```
Task 2.0 Task 2.0 (3 marks), use R to visualize the latitude and longitude data. Specifically create a map that shows where Uber pick-ups occur in NYC. Perform this via points plot on a spatial map (1 plot ¡V 0.5 mark), and via density plots (1 plot ¡V 0.5 mark). 
```{r}
library(ggmap)
#install.packages("ggmap")
#Due to the power of computer, I can only read two data set April and May.
lessUberRawData <- rbind(Data1,Data2)
#use qmap to get the map of New York City
new_york_map <- get_map(location = "New_York")
new_york_map_zoom <- get_map(location = "New_York", zoom=13)
#point plot
ggmap(new_york_map) + geom_point(aes(x=Lon, y=Lat), data=lessUberRawData, col="orange", alpha=0.4, size=0.5)
#density plot
ggmap(new_york_map_zoom) + geom_density_2d(data=lessUberRawData, aes(x=Lon, y=Lat))+stat_density2d(data = lessUberRawData, aes(x=Lon, y=Lat, fill=..level.., alpha=..level..), size=0.01, bins=16, geom="polygon")+scale_fill_gradient(low="green",high="red")+scale_alpha(range=c(0.00,0.25),guide=F)+theme(legend.position = "none", axis.title=element_blank(), text= element_text(size=12))
```
Task 2.1 Show how the number of pick-up vary per day of the week and according to location. 
```{r}
#Set the location to adjust the setting of default language
Sys.setlocale("LC_TIME", "C")
#Make the date and time a formal format
lessUberRawData$Date.Time<- as.POSIXct(lessUberRawData$Date.Time, format="%m/%d/%Y %H:%M:%S")
#Assign the dataof Lan, Lon and weekday(use the as.Date to automatically transform the date to weekday).
weekday_pickup <- data.frame(lessUberRawData$Lat, lessUberRawData$Lon,weekdays(as.Date(lessUberRawData$Date.Time)))
colnames(weekday_pickup) <- c("Lat", "Lon", "Weekdays")
#Re-order the weekdays
weekday_pickup$Weekdays <- ordered(weekday_pickup$Weekdays, levels = c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))
#Facet wrap to print out the plots according to the weekdays.
ggmap(new_york_map_zoom) + geom_density_2d(data=weekday_pickup, aes(x=Lon, y=Lat))+stat_density2d(data = weekday_pickup, aes(x=Lon, y=Lat, fill=..level.., alpha=..level..), size=0.01, bins=16, geom="polygon")+scale_fill_gradient(low="green",high="red")+scale_alpha(range=c(0.00,0.25),guide=F)+theme(legend.position = "none", axis.title=element_blank(), text= element_text(size=12)) + facet_wrap(~Weekdays, nrow=3)+labs(title="Number of pick-up per day of the week", subtitle="From Monday to Sunday")
```
#Task2.2 Show how the number of pick-up vary per month and according to location. 
```{r}
#install.packages("plyr")
library(plyr)
#Due to the power of the computer, I can only read three datasets from April to June
threeUberRawData <- rbind(Data1, Data2, Data3)
#extract the information of Lat, Lon and Month and then store them into a data frame month_pickup
month_pickup <- data.frame(threeUberRawData$Lat, threeUberRawData$Lon, substring(threeUberRawData$Date.Time,1,1))
colnames(month_pickup) <- c("Lat","Lon","Month")
#Re-order by creating a new column called Month_f
month_pickup$Month_f <- factor(month_pickup$Month, levels = c("4","5","6"))
month_pickup$Month_f <- as.character(month_pickup$Month_f)
#Change the name of month from number to word format
month_pickup$Month_f<-revalue(as.character(month_pickup$Month_f), c("4"="April", "5"="May", "6"="June"))
#Plot by facet wrap
ggmap(new_york_map_zoom) + geom_density_2d(data=month_pickup, aes(x=Lon, y=Lat))+stat_density2d(data = month_pickup, aes(x=Lon, y=Lat, fill=..level.., alpha=..level..), size=0.01, bins=16, geom="polygon")+scale_fill_gradient(low="green",high="red")+scale_alpha(range=c(0.00,0.25),guide=F)+theme(legend.position = "none", axis.title=element_blank(), text= element_text(size=12)) + facet_wrap(~Month_f, ncol=3)+labs(title="Number of pick-up for each month ", subtitle="From April to September")
```
Task 2.3 Show how the number of pick-up vary per hour of the day and according to location. 
```{r}
#Due to the power of Computer, I can only read three dataset from April to June
threeUberRawData <- rbind(Data1, Data2, Data3)
#Give a formal format to the date and time and then update them into the column of threeUberRawData
threeUberRawData$Date.Time <- as.POSIXct(threeUberRawData$Date.Time, format="%m/%d/%Y %H:%M:%S")
#Cut the hour part of Date.Time column and store it with lat and lon into hour_pickup
hour_pickup <- data.frame(threeUberRawData$Lat, threeUberRawData$Lon, substring(threeUberRawData$Date.Time,12,13))
colnames(hour_pickup) <- c("Lat","Lon","Hour")
#Re-order by the hour from 00:00 to 23:00
hour_pickup$Hour_f <- factor(hour_pickup$Hour, levels = c("00","01","02","03","04","05","06","07","08","09","10","11","12","13","14","15","16","17","18","19","20","21","22","23"))
#Plot by facet wrap
ggmap(new_york_map_zoom) + geom_density_2d(data=hour_pickup, aes(x=Lon, y=Lat))+stat_density2d(data = hour_pickup, aes(x=Lon, y=Lat, fill=..level.., alpha=..level..), size=0.01, bins=16, geom="polygon")+scale_fill_gradient(low="green",high="red")+scale_alpha(range=c(0.00,0.25),guide=F)+theme(legend.position = "none", axis.title=element_blank(), text= element_text(size=12)) + facet_wrap(~Hour_f, ncol=6)+labs(title="Number of pick-up for each month ", subtitle="From 00:00AM to 23:59PM")
```
Task 3.1 To calculate the cosine similarity between every pair of automobiles if we represent each automobile as a vector of the above 11 variables in ¡§mtcars¡¨. This question requires you to create a matrix (e.g., A) to store the calculation results.
```{r}
#install.packages("lsa")
#install.packages("SnowballC")
library(lsa)
library(SnowballC)
#Use the data form the built-in database
data("mtcars")
#use the cosine function to get the matrix of similarity
similarity_car <- cosine(t(as.matrix(mtcars)), NULL)
#print out the similarity matrix
similarity_car

```

Task 3.2 To find the most similar automobile (only one is OK) for each one and print the results as follows: 
```{r}
#In order to exclude the interference of same car, set all the elements on the diagonal as 0
for(i in 1:ncol(similarity_car)){
  similarity_car[i,i] <- NA
  
}
#initialize a data frame - most_similar_car
most_similar_car <- as.data.frame(matrix(nrow = 0, ncol = 2))
#use for to deal with each column one by one
for(i in 1:ncol(similarity_car)){
  #Input the ith column element into a vector
  vec_car <- c(as.numeric(similarity_car[,i]))
  #missing value will cause the max fun to return NA
  #Find out the max number which mean the highest similarity and return the index
  index <- which(vec_car==max(vec_car, na.rm = T), arr.ind = T)
  #Create the data frame to store the column name and row names of the two cars with highest similarity
  temp <- data.frame(colnames(similarity_car)[i], rownames(similarity_car)[index])
  #Record all the similar car
  most_similar_car <- rbind(most_similar_car, temp)
}
#Print out the list of cars 
most_similar_car
```